Implementing Capon Beamforming on the GPU for Real Time Cardiac Ultrasound Imaging

### 2013-01-25 HRUIonGPU/jrnl_article.tex
uffc

\begin{figure*}[!t]
\centerline{\subfloat[]{\includegraphics[width=3.5in]{gfx/benchmark_bar_bs_M=64_Nx=80_Ny=400_Nb=3.png} \label{fig:benchBS3}}
\hfil
\subfloat[]{\includegraphics[width=3.5in]{gfx/benchmark_bar_bs_M=96_Nx=120_Ny=400_Nb=3.png}%
\label{fig:benchBS5}}}
\caption{Benchmark of GPU beamspace Capon beamforming for a cardiac image covering a $70 \degree$ and $15$cm deep sector. Execution time is plotted as a function of subarray length $L$.  If $L$ is kept constant on $M/2$ and execution time is plotted against $N_b$, the plot would be similar to the two in Fig. \ref{fig:bench}. (a) 64 elements array. (b) 96 elements array.}
\label{fig:benchBS}
\end{figure*}
In this section we present benchmarks of the presented implementations performed on a Nvidia Quadro 6000 graphics card (14 SMs, 1 TFlops theoretical single precision, 6 GB global memory, and 144 GB/s memory bandwidth). Problem sizes are calculated based on 32, 64, and 96 element phased arrays with a 2.5 MHz center frequency. Assuming a system bandwidth of 80\% we need a complex sampling frequency of 2 MHz, and when imaging down to a depth of 15 cm we end up with approximately 400 samples per range line. To correctly sample a $70\degree$ sector with 32, 64 and 96 element arrays ($\lambda/2$ pitch), we need at least 40, 80 and 120 beams respectively. The presented execution times do not include transfer of data from CPU-side to the GPU, since data are assumed to be pre-delayed and available in GPU memory. What is included are all calculations, memory transfers from global to near-core GPU memory, and near core memory access.

A benchmark of the implementation described in Section \ref{sec:meth} is presented in Fig. \ref{fig:bench}. Benchmarks are based on 32 and 64 elements phased arrays respectively, using 40 and 80 transmit beams. For each subarray size the total execution time is split into one bar for each of the three steps described in Section \ref{sec:meth}, and calculation of $\mat{\hat{R}}$ is presented for three different values of $K$. For $M=64$, $L=32$ and $K=2$ (Fig. \ref{fig:benchUS64}), a typical setup for high quality cardiac imaging, processing throughput is 10 frames per second. In Fig. \ref{fig:benchBS} we present similar benchmarks for the beamspace version of the Capon beamformer introduced in Section \ref{sec:bs}. Benchmarks are based on 64 and 96 elements phased arrays, using 80 and 120 transmit beams. The figure only present results with use of temporal averaging ($K=2$). For $M=96$, $L=48$, $K=2$ and $N_b=3$, processing throughput is 100 frames per second. For all benchmarks there is a line indicating the time it takes to acquire the underlying image using single line acquisition. Hence, below this line we are capable of real time processing.

\section{Simulated phantom and \textit{in-vivo} cardiac data}

@@

Fig. \ref{fig:invivo} presents results of applying Capon and BS Capon on an \textit{in-vivo} image of the left ventricle. The image was acquired using a experimental research scanner at MI-Lab in Trondheim using a 3.5 MHz phased array in harmonic mode. The image is extracted from a cardiac cycle, that can be viewed in its full length in an attached video. Its purpose is to show that Capon beamforming in both element- and beam-space is feasible for cardiac imaging, not just in simulations.

To get a better impression of the high resolution images obtained with Capon beamforming, two slices taken from the images in Fig. \ref{fig:phantom} and \ref{fig:invivo} is included in Fig. \ref{fig:slices}.

Both data sets have been processed with the following Capon beamformer settings: $L = 32$, $K = 2$, $d=0.1$. These settings have previously been shown to provide high resolution while maintaining DAS-like speckle\cite{Synnevag2007a}. For the beamspace processing the following parameters have been used: $L=32$, $K=2$, $d=0.01$, $N_b=3$. For the simulation, 16x lateral over-sampling has been used, in order to avoid under sampling of the two bright point scatterers as they move. The cardiac recording only has $30\%$ lateral over sampling.

\section{Trading resolution for speed}
As we see from Fig. \ref{fig:bench} and \ref{fig:benchBS}, execution time is reduced when either $L$ or $N_b$ is reduced. The question is how does this impact image quality. In Fig. \ref{fig:speed_res_trade} minimum distance between the two point targets in Fig. \ref{fig:phantom} (while preserving a 6dB lower level in between) is plotted against $L$ (when processed with Capon) and $N_b$ (when processed with BS Capon). For high values the two methods are equal, but for lower values BS Capon provides higher resolution than what is possible with small subarrays. For this result, $d=0.01$ has been used on both methods. Why resolution decreases for BS Capon for small $N_b$s is partially explained by reduced information of interference at higher angles, and automatically increased diagonal loading. It is also interesting to see that Capon beamforming can perform worse than DAS for small subarrays.

@@

\section{Discussion}\label{sec:dis}
We have presented a parallel implementation of both the Capon og BS Capon beamformer.

@@

Toeplitz-based solver (cite new article). Discuss it.

@@

>>>>

\begin{figure*}[!t]
\centerline{\subfloat[]{\includegraphics[width=3.5in]{gfx/benchmark_bar_bs_M=64_Nx=80_Ny=400_Nb=3.png} \label{fig:benchBS3}}
\hfil
\subfloat[]{\includegraphics[width=3.5in]{gfx/benchmark_bar_bs_M=96_Nx=120_Ny=400_Nb=3.png}%
\label{fig:benchBS5}}}
\caption{Benchmark of GPU beamspace Capon beamforming for a cardiac image covering a $70 \degree$ and $15$cm deep sector. Execution time is plotted as a function of subarray length $L$.  If $L$ is kept constant on $M/2$ and execution time is plotted against $N_b$, the plot would be similar to the two in Fig. \ref{fig:bench}. (a) 64 elements array. (b) 96 elements array.}
\label{fig:benchBS}
\end{figure*}
In this section we present benchmarks of the implementations presented in Section \ref{sec:meth} and \ref{sec:bs} performed on a Nvidia Quadro 6000 graphics card (14 SMs, 1 TFlops theoretical single precision, 6 GB global memory, and 144 GB/s memory bandwidth). Problem sizes are calculated based on 32, 64, and 96 element phased arrays with a 2.5 MHz center frequency. Assuming a system bandwidth of 80\% we need a complex sampling frequency of 2 MHz, and when imaging down to a depth of 15 cm we end up with approximately 400 samples per range line. To correctly sample a $70\degree$ sector with 32, 64 and 96 element arrays ($\lambda/2$ pitch), we need at least 40, 80 and 120 beams respectively. The presented execution times do not include transfer of data from CPU-side to the GPU, since data are assumed to be pre-delayed and available in GPU memory. What is included are all calculations, memory transfers from global to near-core GPU memory, and near core memory access.

A benchmark of the implementation described in Section \ref{sec:meth} is presented in Fig. \ref{fig:bench}. Benchmarks are based on 32 and 64 elements phased arrays respectively, using 40 and 80 transmit beams. For each subarray size the total execution time is split into one bar for each of the three steps described in Section \ref{sec:meth}, and calculation of $\mat{\hat{R}}$ is presented for three different values of $K$. For $M=64$, $L=32$ and $K=2$ (Fig. \ref{fig:benchUS64}), a typical setup for high quality cardiac imaging, processing throughput is 10 frames per second.

In Fig. \ref{fig:benchBS} we present similar benchmarks for the beamspace version of the Capon beamformer introduced in Section \ref{sec:bs}. Benchmarks are based on 64 and 96 elements phased arrays, using 80 and 120 transmit beams. The figure only present results with use of temporal averaging ($K=2$). For $M=96$, $L=48$, $K=2$ and $N_b=3$, processing throughput is 100 frames per second. For all four benchmarks there is a line indicating the time it takes to acquire the underlying image using single line acquisition. Hence, below this line we are capable of real time processing when a single receive line is used per transmit line.

\section{Simulated phantom and \textit{in-vivo} cardiac images}

@@

Fig. \ref{fig:invivo} presents results of applying Capon and BS Capon on an \textit{in-vivo} image of the left ventricle. The image was acquired using a experimental research scanner at MI-Lab in Trondheim using a 64 elements, 3.5 MHz, and $\lambda/2$ pitched phased array in harmonic mode. The image is extracted from a cardiac cycle, that can be viewed in its full length in an attached video. Its is included to show that Capon beamforming in both element- and beam-space is feasible for cardiac imaging, not just in simulations.

To get a better impression of the high resolution images obtained with Capon beamforming, two slices taken from the images in Fig. \ref{fig:phantom} and \ref{fig:invivo} is included in Fig. \ref{fig:slices}. For the phantom, the cut goes through the two point targets at 93 mm depth, and for the cardiac image it goes through the tip of the mitral valve at 96 mm depth.

Both data sets have been processed with the following Capon beamformer settings: $L = 32$, $K = 2$, $d=0.1$. These settings have previously been shown to provide high resolution while maintaining DAS-like speckle\cite{Synnevag2007a}. For the beamspace processing the following parameters have been used: $L=32$, $K=2$, $d=0.01$, $N_b=3$. For the simulated phantom, 16x lateral over-sampling has been used, in order to avoid under sampling of the two bright point scatterers as they move. The cardiac recording only has $30\%$ lateral over sampling.

\section{Trading resolution for speed}
As we see from Fig. \ref{fig:bench} and \ref{fig:benchBS}, execution time is reduced when either $L$ or $N_b$ is reduced. So to gain an  increase in processing throughput we can reduce $L$ or $N_b$. The question is how does this impact image quality. In Fig. \ref{fig:speed_res_trade} minimum distance between the two point targets in Fig. \ref{fig:phantom} (while preserving a 6dB lower level in between) is plotted against $L$ (when processed with Capon) and $N_b$ (when processed with BS Capon). For high values the two methods are equal, but for lower values, BS Capon provides higher resolution than what is possible with small subarrays. For this result, $d=0.01$ has been used on both methods. Why resolution decreases for BS Capon for small $N_b$s is partially explained by reduced information of interference at higher angles, and an automatically increase in diagonal loading. It is also interesting to see that Capon beamforming performs worse than DAS for small subarrays. (Comment on beamspace).

@@

\section{Discussion}\label{sec:dis}
In this paper we have presented a parallel implementation of the Capon beamformer both in element- and beam-space. The processing rates acquired shows that it is now possible to apply BS Capon in real time in cardiac ultrasound imaging. From Fig. ... 2MLA is supported.

@@

@@

MLA's has, to some extent be used to increase later sampling. When using a high resolution beamformer, Nyquist tells us that the sampling frequency must be increase. Future work will concentrate on how much over sampling is required when using the Capon beamformer whit i curtain set of parameters.
